\chapter{RL Circuits}
\section{Why RL Circuits Matter}
Inductors appear in RF chokes, tuned circuits, filters, relays, transformers,
and antenna loading coils. An RL circuit is the current-domain sibling of the RC
circuit.

\section{Transient Response}
For a DC step \(E\) applied to a series RL circuit,
\[
i(t)=\frac{E}{R}\left(1-\ee^{-Rt/L}\right).
\]
For decay from initial current \(I_0\),
\[
i(t)=I_0\ee^{-Rt/L}.
\]
The time constant is
\[
\tau=\frac{L}{R}.
\]
Inductor current cannot jump instantaneously under finite voltage.

\section{State-Space Interpretation}
With \(x=i\),
\[
\dot{x}=-\frac{R}{L}x+\frac{1}{L}u.
\]
The eigenvalue is \(-R/L\), and the time constant is \(L/R\).

\section{Frequency Interpretation}
\[
Z_L=\jj\omega L.
\]
At high frequency the inductor's impedance is large, so inductors tend to oppose
rapid changes in current. At DC steady state an ideal inductor behaves like a
short circuit.

\section{Energy Interpretation}
\[
E_L=\frac{1}{2}Li^2.
\]
The magnetic field stores energy when current flows. When current decays, that
energy must go somewhere, which is why interrupting inductive current can create
large voltages.

\begin{exambox}
Remember the natural state: capacitor voltage and inductor current. That one
sentence explains which quantity is continuous and which differential equation
you should write.
\end{exambox}

\section{Worked Example}
A \(\SI{10}{\micro\henry}\) inductor is in series with a \(\SI{50}{\ohm}\)
resistor. Then
\[
\tau=\frac{L}{R}=\frac{10\times10^{-6}}{50}
=\SI{2.0e-7}{\second}.
\]
At \(\SI{14}{\mega\hertz}\),
\[
X_L=2\pi(14\times10^6)(10\times10^{-6})\approx \SI{880}{\ohm}.
\]
