\chapter{Filters, Matching Networks, and Transformers}
\label{ch:filters}
\section{Why Filters Matter in Amateur Radio}
Filters select desired signals and reject unwanted ones. They appear in receiver
front ends, IF stages, audio stages, transmitter outputs, anti-aliasing paths,
power supplies, and interference cures. With the s-plane chapter in hand, a filter
is simply a deliberate placement of poles (to shape the passband) and zeros (to
notch the stopband).

\section{Basic Filter Types}
\begin{description}
\item[Low-pass] Passes low frequencies and attenuates high frequencies.
\item[High-pass] Passes high frequencies and attenuates low frequencies.
\item[Band-pass] Passes a frequency band.
\item[Band-stop or notch] Rejects a frequency band.
\end{description}

\begin{figure}[htbp]
\centering
\begin{circuitikz}
\draw (0,0) node[left]{in} to[R,l=\(R\)] (3,0)
      to[C,l=\(C\)] (3,-2) node[ground]{};
\draw (3,0) to[short,-o] (4,0) node[right]{out};
\end{circuitikz}
\caption{First-order RC low-pass filter.}
\end{figure}

\section{Cutoff, Slope, and Order}
For a first-order RC filter, \(f_c=1/(2\pi RC)\). Past its corner, each pole
changes the asymptotic magnitude by about \(-\SI{20}{\decibel}\) per decade, and
each zero by \(+\SI{20}{\decibel}\) per decade. The \emph{order} of a filter is
its number of poles, so an \(n\)-pole low-pass rolls off at about
\(-20n\,\si{\decibel}\) per decade far above cutoff. Steeper skirts require more
poles.

\section{The Second-Order Section}
Higher-order filters are built from first-order and second-order sections. The
generic second-order low-pass transfer function is
\[
H(s)=\frac{\wzero^2}{s^2+\dfrac{\wzero}{Q}\,s+\wzero^2},
\]
identical in form to the series-RLC response of \cref{ch:splane}. Here the filter
\(Q\) is a design choice, not just a loss parameter: it sets how close the pole
pair sits to the \(\jj\omega\) axis and therefore how much the response peaks near
\(\wzero\). \(Q=1/\sqrt{2}\approx0.707\) gives the maximally flat (Butterworth)
second-order response with no peaking; higher \(Q\) peaks; lower \(Q\) rounds off
early. Band-pass and high-pass sections share the same denominator with different
numerators (a zero at the origin, or a double zero at the origin, respectively).

\section{Filter Families as Pole Patterns}
The named families differ only in \emph{where} the poles (and zeros) are placed
for a given order.

\begin{tabularx}{\textwidth}{@{}L{0.19\textwidth}Y@{}}
\toprule
Family & Pole/zero placement and consequence \\
\midrule
Butterworth & Poles equally spaced on a semicircle of radius \(\wzero\) in the left half-plane. Maximally flat passband, no ripple, moderate transition. \\
Chebyshev & Poles on an ellipse pushed closer to the \(\jj\omega\) axis. Sharper cutoff at the price of passband ripple and worse phase/transient behavior. \\
Elliptic (Cauer; ``elliptical'' in the pool) & Chebyshev-like poles plus transmission \emph{zeros} on the \(\jj\omega\) axis in the stopband. Sharpest transition; ripple in both bands. \\
Bessel & Poles positioned for maximally flat \emph{group delay} (linear phase); gentle amplitude cutoff. Good for pulse fidelity. \\
Crystal / cavity & Very high-\(Q\) resonators (poles hugging the axis) for narrow, highly selective band-pass shapes. \\
DSP FIR & No feedback, so no poles away from the origin (all zeros); can be exactly linear phase. \\
DSP IIR & Recursive: has poles, so it mimics analog families efficiently but with nonlinear phase. \\
\bottomrule
\end{tabularx}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[>=Latex,scale=1.15]
\draw[->,black!70] (-2.7,0) -- (0.8,0) node[right]{\(\sigma\)};
\draw[->,black!70] (0,-2.0) -- (0,2.0) node[above]{\(\jj\omega\)};
\draw[GuideLine,dashed] (0,0) circle (1.6);
% 4th-order Butterworth poles: angles 112.5,157.5,202.5,247.5 deg
\foreach \a in {112.5,157.5,202.5,247.5}{
  \filldraw[GuideBlue] ({1.6*cos(\a)},{1.6*sin(\a)}) circle (2.2pt);
}
\node[GuideLine] at (1.15,1.35) {\small \(|s|=\wzero\)};
\node[GuideBlue] at (-2.15,1.2) {\small Butterworth};
\end{tikzpicture}
\caption{Fourth-order Butterworth poles: four left-half-plane poles equally
spaced on the circle of radius \(\wzero\). Higher order adds more poles on the
same circle; Chebyshev pulls them toward the \(\jj\omega\) axis for a sharper but
rippled response.}
\label{fig:butterworth-poles}
\end{figure}

\begin{controlsbox}
\textbf{Refresher: reading a filter as poles.} Filter design is pole placement in
reverse---you choose a target magnitude response and solve for pole locations.
Butterworth = poles on a circle (flattest); Chebyshev = poles on an ellipse
nearer the axis (higher per-stage \(Q\), sharper, rippled); elliptic adds
\(\jj\omega\)-axis zeros for stopband notches. The trade among flatness,
sharpness, and phase linearity is entirely a trade about how close the poles sit
to the imaginary axis.
\end{controlsbox}

\section{Impedance Matching}
An impedance-matching network does two jobs:
\begin{enumerate}
\item cancel the reactive part of a complex load, and
\item transform the remaining resistance to the desired system impedance.
\end{enumerate}
In a \(\SI{50}{\ohm}\) RF system, matching maximizes power transfer and keeps
transmission-line reflections controlled.

\subsection{The L-Network}
The simplest matching network uses two reactances. To match two real resistances
\(R_{\text{hi}}>R_{\text{lo}}\), place a series reactance on the low-resistance
side and a shunt reactance on the high-resistance side. The network operating
\(Q\) is set by the resistance ratio:
\[
Q=\sqrt{\frac{R_{\text{hi}}}{R_{\text{lo}}}-1},
\qquad
X_{\text{series}}=Q\,R_{\text{lo}},
\qquad
X_{\text{shunt}}=\frac{R_{\text{hi}}}{Q}.
\]
The two reactances are of opposite type (one inductor, one capacitor); choosing
which is the inductor sets whether the network is low-pass or high-pass. An
L-network has no free parameter left over, so its bandwidth is fixed once the
resistances are chosen.

\subsection{Pi and T Networks}
Adding a third element gives a \(\pi\) or T network. The extra element lets the
designer choose the loaded \(Q\)---and therefore the bandwidth and harmonic
rejection---independently of the resistance ratio. This is why transmitter output
and antenna-tuner networks are usually \(\pi\) or T: they match \emph{and} set
selectivity. Higher loaded \(Q\) narrows bandwidth and improves harmonic
suppression at the cost of higher circulating currents and losses.

\begin{mistakebox}
Matching is not the same thing as filtering. A network can do both, but the
concepts are distinct: matching controls the impedance seen by a source or line,
while filtering controls frequency-dependent transmission. A \(\pi\)-network tuner
happens to be a low-pass filter too, which is a feature, not the definition.
\end{mistakebox}

\section{Transformers}
For an ideal transformer,
\[
\frac{V_s}{V_p}=\frac{N_s}{N_p},
\qquad
\frac{I_s}{I_p}=\frac{N_p}{N_s},
\qquad
Z_p=\left(\frac{N_p}{N_s}\right)^2 Z_s.
\]
Impedance transforms as the turns ratio \emph{squared}. Core saturation should be
avoided: a saturated core can no longer support proportional flux change, causing
distortion, heating, and loss of the intended impedance transformation.

\section{Worked Example: Quarter-Wave and L-Network}
To match a real \(\SI{200}{\ohm}\) load to a \(\SI{50}{\ohm}\) system with a
lossless quarter-wave transformer,
\[
Z_t=\sqrt{Z_0Z_L}=\sqrt{(50)(200)}=\SI{100}{\ohm}.
\]
The same match with an L-network gives
\[
Q=\sqrt{\frac{200}{50}-1}=\sqrt{3}\approx1.73,
\qquad
X_{\text{series}}=1.73\times50\approx\SI{87}{\ohm},
\qquad
X_{\text{shunt}}=\frac{200}{1.73}\approx\SI{116}{\ohm}.
\]
Both realize the match; the quarter-wave version needs a line one quarter
wavelength long at the design frequency, while the L-network is lumped and
narrower band.
