\chapter{Resistive Networks}
\label{ch:resistive}
Resistive networks set bias points, divide voltages, sample currents, terminate
transmission lines, and model losses. They are the simplest place to practice the
conservation laws of \cref{ch:foundations}---KCL and KVL---and the equivalent-circuit
idea, so every rule below is derived from those laws rather than memorized.

\section{Series and Parallel from KCL and KVL}
\textbf{Series} resistors carry the \emph{same current} \(I\). KVL sums their
voltages, so
\[
V=IR_1+IR_2+\cdots=I\sum_k R_k
\quad\Longrightarrow\quad
R_{\mathrm{eq}}=\sum_k R_k .
\]
\textbf{Parallel} resistors share the \emph{same voltage} \(V\). KCL sums their
currents, so
\[
I=\frac{V}{R_1}+\frac{V}{R_2}+\cdots=V\sum_k\frac{1}{R_k}
\quad\Longrightarrow\quad
\frac{1}{R_{\mathrm{eq}}}=\sum_k\frac{1}{R_k},
\qquad
R_{\mathrm{eq}}=\frac{R_1R_2}{R_1+R_2}\ \text{(two resistors)}.
\]
Series adds resistances; parallel adds conductances. That single distinction---same
current vs.\ same voltage---is the whole idea.

\section{Voltage and Current Dividers}
A voltage divider is just two series resistors read across one of them
(\cref{fig:divider}). The same current \(I=V_i/(R_1+R_2)\) flows through both, so
the output taken across \(R_2\) is
\[
V_o=IR_2=V_i\frac{R_2}{R_1+R_2}.
\]

\begin{figure}[htbp]
\centering
\begin{circuitikz}[scale=1]
\draw (0,0) to[V,l=\(V_i\)] (0,3)
      to[R,l=\(R_1\)] (2.5,3)
      to[R,l=\(R_2\),v=\(V_o\)] (2.5,0) -- (0,0);
\draw (2.5,3) to[short,-o] (3.5,3) node[right]{$+$};
\draw (2.5,0) to[short,-o] (3.5,0) node[right]{$-$};
\end{circuitikz}
\caption{Voltage divider: the output across \(R_2\) is
\(V_o=V_i R_2/(R_1+R_2)\).}
\label{fig:divider}
\end{figure}

Dually, current entering two parallel resistors splits \emph{inversely} with
resistance,
\[
I_1=I\frac{R_2}{R_1+R_2},\qquad I_2=I\frac{R_1}{R_1+R_2},
\]
so more current takes the lower-resistance branch.

\section{Thevenin and Norton Equivalents}
Any linear one-port of sources and resistors looks, from its terminals, like a
single Thevenin source \(V_{Th}\) in series with \(R_{Th}\), or equivalently a
Norton source \(I_N\) in parallel with \(R_N\) (\cref{fig:thevenin}):
\[
R_N=R_{Th},\qquad I_N=\frac{V_{Th}}{R_{Th}}.
\]
This matters in radio because a receiver input, an antenna, an amplifier stage, or
a test instrument is usually characterized by exactly this terminal
model---one source and one impedance---so you can predict how it behaves when a
different load or stage is connected without re-analyzing the whole circuit.

\begin{figure}[htbp]
\centering
\begin{circuitikz}
\draw (0,0) to[V,l=\(V_{Th}\)] (0,2)
      to[R,l=\(R_{Th}\)] (3,2)
      to[short,-o] (4,2);
\draw (0,0) -- (3,0) to[short,-o] (4,0);
\node at (4.8,1) {\(\Longleftrightarrow\)};
\draw (6,0) -- (6,2);
\draw (6,0) to[I,l_=\(I_N\)] (6,2);
\draw (6,2) to[R,l=\(R_N\)] (9,2) -- (9,0) -- (6,0);
\draw (9,2) to[short,-o] (10,2);
\draw (9,0) to[short,-o] (10,0);
\end{circuitikz}
\caption{Equivalent Thevenin and Norton one-port descriptions.}
\label{fig:thevenin}
\end{figure}

\section{Power Transfer and Matching}
For a purely resistive Thevenin source, maximum power reaches the load when
\[
R_L=R_{Th},
\]
and for complex impedances the maximum-real-power condition is the conjugate match
\(Z_L=Z_S^{*}\). This is why RF systems are built around a common reference
impedance: matching the source, line, load, and instruments to (typically)
\(\SI{50}{\ohm}\) delivers the most power to the antenna and, as \cref{ch:lines}
shows, minimizes reflections on the feed line.

\begin{exambox}
\(\SI{50}{\ohm}\) is a design convention, not a law of nature. Matching means the
system has been arranged so sources, lines, loads, and measuring instruments
interact predictably---maximum power transfer and minimum reflection---not that
every impedance is \(\SI{50}{\ohm}\).
\end{exambox}

\section{Worked Example}
A \(\SI{12}{\volt}\) source feeds \(R_1=\SI{2.2}{\kilo\ohm}\) on top and
\(R_2=\SI{1.0}{\kilo\ohm}\) to ground (the topology of \cref{fig:divider}). The
unloaded output and the Thevenin resistance seen by a load are
\[
V_o=12\frac{1.0}{2.2+1.0}=\SI{3.75}{\volt},
\qquad
R_{Th}=R_1\parallel R_2=\frac{(2.2)(1.0)}{2.2+1.0}\,\si{\kilo\ohm}
\approx\SI{0.688}{\kilo\ohm}.
\]
So to any load, this divider looks like a \(\SI{3.75}{\volt}\) source behind
\(\SI{688}{\ohm}\)---and a load comparable to \(\SI{688}{\ohm}\) will noticeably
pull the output down, which is why divider outputs are buffered before driving a
real load.
