\chapter{Measurement and Troubleshooting}
\section{Why Measurement Is a Circuit Problem}
Meters, probes, oscilloscopes, antenna analyzers, and SWR bridges are themselves
circuits. Connecting a measurement instrument changes the circuit unless the
instrument impedance is high, low, or matched in the right way for the
measurement.

\section{Voltmeter and Ammeter Loading}
A voltmeter should usually have high input impedance so it draws little current.
An ammeter should have low series resistance so it introduces little voltage
drop. At RF, even a short lead can add inductance or capacitance.

\section{Oscilloscope and Probe Ideas}
A scope probe has resistance and capacitance. A compensated probe is adjusted so
the divider ratio is approximately independent of frequency over the intended
range. Ground-lead inductance can distort fast or RF measurements.

\section{SWR Bridge and Antenna Analyzer}
An SWR bridge estimates forward and reflected wave components on a transmission
line. An antenna analyzer can measure impedance and SWR versus frequency, making
it useful for finding resonance and matching conditions.

\section{Decibels in Measurement}
\[
P_{\dB}=10\log_{10}\frac{P}{P_{\mathrm{ref}}},
\qquad
V_{\dB}=20\log_{10}\frac{V}{V_{\mathrm{ref}}}.
\]
Always check the reference. dBW references \(\SI{1}{\watt}\); dBm references
\(\SI{1}{\milli\watt}\).

\section{Troubleshooting Pattern}
\begin{enumerate}
\item Identify the expected DC operating point.
\item Check whether the signal path is AC-coupled, DC-coupled, filtered, or
matched.
\item Estimate impedance loading before attaching an instrument.
\item Compare measured gain, attenuation, phase, or SWR against the simple
model.
\item If results differ, ask which assumption failed: ideal element, lumped
model, steady state, no loading, no parasitics, or no nonlinear behavior.
\end{enumerate}

\begin{exambox}
Measurement questions often reward recognizing what the instrument actually
loads or reports: voltage, current, power, impedance, SWR, frequency, or phase.
\end{exambox}
