\chapter*{How to Use This Book}
\addcontentsline{toc}{chapter}{How to Use This Book}
The book is organized in four parts, and how you read them depends on your goal:
\begin{description}
\item[Part I --- Exam-Ready Reference] a compact cheat sheet. If the exam is
close and you only need a formula refresher, this part (with the practice in
Part~IV) is enough.
\item[Part II --- Mathematical and Control-Theory Foundations] the generic
toolkit: complex numbers and phasors; first- and second-order linear dynamical
systems (Laplace, block diagrams, feedback); poles, zeros, and the s-plane; and
Bode plots. Read this if you want to \emph{understand} rather than memorize.
\item[Part III --- The Circuits] each amateur-radio circuit built up as an
instance of the Part~II toolkit. This is where the understanding pays off.
\item[Part IV --- Practice and Study] worked examples and reasoning through real
question-pool items.
\end{description}

The engineering path is to read Part~II, then Part~III. Each circuit chapter in
Part~III follows the same recipe:
\begin{enumerate}
\item draw the schematic and name the energy-storage state variables;
\item apply KCL/KVL and the element laws to get the differential equation;
\item write the state-space model and read off the poles;
\item match it to the generic first- or second-order system of Part~II;
\item interpret the frequency response (Bode) and pole motion (root locus);
\item close with the energy balance and a worked example.
\end{enumerate}

Throughout, colored sidebars separate the layers: \emph{Exam Relevance} and
\emph{Common Mistake} for exam-level facts, and \emph{Controls Connection},
\emph{Physical Insight}, and \emph{Energy Insight} for the deeper interpretation.
For a fast review, read the \emph{Exam Relevance} and \emph{Common Mistake} boxes
first.
