\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
This book teaches the circuit theory behind the Amateur Extra exam from the
perspective of a control engineer. The intended reader already knows differential
equations, eigenvalues, and state-space systems, and has at least a rusty
acquaintance with classical frequency-domain ideas---poles and zeros, Bode plots,
and the root locus. A chemical, mechanical, electrical, or systems engineer
returning to circuits after years away is exactly the audience. Such a reader does
not want to memorize exam answers; they want to \emph{see} each radio circuit as a
dynamical system: to write its differential equation, cast it in state-space form,
locate its poles, and read its Bode and root-locus plots. The organizing idea of
the book is that one circuit can be understood as a physical object, a set of
conservation laws, a differential equation, a state-space system, a
phasor/frequency response, and an energy balance---and that these views must
agree.

The book is in three parts. Part~I is a compact, exam-ready reference. Part~II
reviews, generically, the mathematics and control theory the rest of the book
leans on: complex numbers and phasors, first- and second-order linear dynamical
systems, poles, zeros, and the s-plane, and frequency response. Part~III then
builds each amateur-radio circuit up as an instance of that generic machinery,
one circuit at a time. Laplace transforms and state-space models appear
throughout as bridges to familiar engineering language; they are not calculations
the Amateur Extra exam directly asks the reader to perform.

This is an independent educational study guide. It is not affiliated with,
endorsed by, or sponsored by the ARRL, NCVEC, or the FCC. Its worked practice is
anchored to the current (2024--2028) Amateur Extra question pool, but question
identifiers and wording change each cycle, so final preparation should be checked
against the official current pool and the current FCC rules.

Several references informed the book's coverage and terminology: the \emph{ARRL
Extra Class License Manual}, the NCVEC Element~4 question pool as published at
HamExam.org, and earlier study guides including Dan Romanchik's (KB6NU) and
others. These were used only as checks for topic coverage, terminology, and
formulas; their prose, questions, answer choices, and figures were not copied.
Full citations appear in the Bibliography.
