\chapter{Formula Index}
\begin{longtable}{@{}L{0.34\textwidth}L{0.58\textwidth}@{}}
\toprule
Formula & Use \\
\midrule
\endhead
\(i=\dd q/\dd t\) & Current as rate of charge flow. \\
\(V=IR\) & Resistor law. \\
\(P=VI=I^2R=V^2/R\) & Power in resistive circuits. \\
\(i=C\,\dd v/\dd t\) & Capacitor current-voltage law. \\
\(v=L\,\dd i/\dd t\) & Inductor voltage-current law. \\
\(E_C=\frac{1}{2}CV^2\) & Capacitor energy. \\
\(E_L=\frac{1}{2}LI^2\) & Inductor energy. \\
\(\tau_{RC}=RC\) & RC time constant. \\
\(\tau_{RL}=L/R\) & RL time constant. \\
\(Z_R=R\) & Resistor impedance. \\
\(Z_L=\jj\omega L\) & Inductor impedance. \\
\(Z_C=1/(\jj\omega C)\) & Capacitor impedance. \\
\(\omega=2\pi f\) & Angular frequency conversion. \\
\(\omega_0=1/\sqrt{LC}\) & Undamped LC angular frequency. \\
\(f_0=1/(2\pi\sqrt{LC})\) & Resonant frequency in hertz. \\
\(Q_s=\omega_0L/R\) & Series RLC quality factor. \\
\(BW\approx f_0/Q\) & Narrowband half-power bandwidth estimate. \\
\(s^2+2\zeta\omega_0 s+\omega_0^2\) & Standard second-order pole polynomial. \\
\(\alpha=R/2L,\ \zeta=\alpha/\omega_0\) & Series RLC damping rate and ratio. \\
\(s=-\alpha\pm\jj\omega_d,\ \omega_d=\sqrt{\omega_0^2-\alpha^2}\) & Underdamped pole pair. \\
\(Q=1/(2\zeta)\) & Quality factor and damping ratio. \\
\(S=VI^*\) & Complex power using RMS phasors. \\
\(A_{\mathrm{cl}}=A/(1+A\beta)\) & Closed-loop gain; \(L=A\beta\) is loop gain. \\
\(\text{gain}\times\text{BW}=A_0\omega_a\) & Gain--bandwidth product (single-pole amp). \\
\(|L|=1,\ \angle L=0\) & Barkhausen oscillation condition. \\
\(Q=\sqrt{R_{\text{hi}}/R_{\text{lo}}-1}\) & L-network matching \(Q\). \\
\(Z_0=\sqrt{L'/C'},\ v_p=1/\sqrt{L'C'}\) & Line impedance and phase velocity. \\
\(\Gamma=(Z_L-Z_0)/(Z_L+Z_0)\) & Load reflection coefficient. \\
\(\mathrm{SWR}=(1+|\Gamma|)/(1-|\Gamma|)\) & Standing wave ratio. \\
\(RL=-20\log_{10}|\Gamma|\) & Return loss in decibels. \\
\(Z_{\mathrm{in}}=Z_0\dfrac{Z_L+\jj Z_0\tan\beta\ell}{Z_0+\jj Z_L\tan\beta\ell}\) & Input impedance along a lossless line. \\
\(Z_t=\sqrt{Z_0Z_L}\) & Quarter-wave transformer for real impedances. \\
\bottomrule
\end{longtable}
