\chapter{Glossary}
\begin{description}
\item[Admittance] Reciprocal impedance, \(Y=1/Z\), measured in siemens.
\item[Bandwidth] Frequency span over which a circuit meets a stated response criterion.
\item[Bode plot] Magnitude and phase of a frequency response versus logarithmic frequency.
\item[Characteristic impedance] Traveling-wave voltage/current ratio on a transmission line.
\item[Corner frequency] Frequency where a first-order response begins its asymptotic slope change.
\item[Cutoff frequency] Often the \(-\SI{3}{\decibel}\) frequency for a filter.
\item[Damping ratio] Dimensionless \(\zeta\) grading a second-order system; \(\zeta=1/(2Q)\).
\item[Gain--bandwidth product] Constant product of closed-loop gain and bandwidth for a single-pole amplifier.
\item[Impedance] Complex ratio \(Z=V/I\) for sinusoidal steady state.
\item[Loaded Q] Quality factor including external loading or coupling.
\item[Loop gain] Gain around an opened feedback loop, \(L=A\beta\); sets stability and desensitization.
\item[Phase margin] Amount by which loop phase exceeds \(-180^\circ\) at unity loop gain.
\item[Phasor] Complex representation of sinusoidal magnitude and phase.
\item[Pole] Root of a transfer-function denominator; equivalently an eigenvalue and a natural mode.
\item[Reactance] Imaginary part of impedance associated with energy storage.
\item[Root locus] Path traced by the poles as a parameter (e.g.\ \(R\) or feedback) varies.
\item[Susceptance] Imaginary part of admittance, \(B\); for a pure reactance \(B=-1/X\).
\item[Reflection coefficient] Complex ratio of reflected to incident wave amplitude.
\item[Return loss] Decibel measure of reflection magnitude.
\item[State variable] Variable that captures stored energy needed to predict future behavior.
\item[Unloaded Q] Quality factor due to internal resonator loss alone.
\item[Zero] Root of a transfer-function numerator; a frequency of null transmission (e.g.\ a trap).
\end{description}
